Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-04 snTomas.tex
incorporated most comments from JH,KS, others

\change{
One of the key values in observing standard candles behind
gravitational lenses is that most lensing reconstructions suffer from
the mass-sheet degeneracy \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of mass in front of or behind the lens, without
disturbing the primary observable quantities.  For example, take a
lens model with a given surface mass density $\kappa$, and then
transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.
}

@@

As described below, we fit the SN \tomas\ spectrum in two steps.
First we determine a spectral classification -- and get a preliminary
estimate of the redshift and age -- using the SuperNova IDentification
(SNID) software \citep{Blondin:2007}.  Second, we refine the redshift
and age measurement using a custom Type Ia spectral template matching
program.

@@

As in \P14, we derive a distance modulus\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.}  from the SALT2 fit using

@@

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  \change{Four of the tested models
(CATS, GLAFIC, Zitrin-NFW, Merten) report a median value that falls
within the 1$\sigma$ uncertainty range of the measured magnification,
and two of the models (Williams, Merten) report uncertainties that
overlap the central $\mu$ from at least one of the light curve
fitters.}  Furthermore, the scatter amongst
models provides a reasonable estimate of the magnification
uncertainty.

@@

\change{
\subsection{Misidentification of System 3}
\label{sec:MisidentificationOfSystem3}
}
\change{
It is important to note that in regards to the impact on this
particular sight-line, not all strong-lensing constraints are equal.
An error in the association of multiple images or in the redshift of a
multiply-imaged galaxy would have a more significant impact if that
system has a projected position close to our SN sight-line.  It
happens that there is one such system, for which the position of a
multiple image near SN \tomas\ is presently disputed.  }

@@

We have speculated on the origin of this systematic bias,
evaluating five possible causes:

@@

@@

Increasing the sample of SNe behind well studied strong lensing
clusters would allow this test to be repeated and refined.  With 10 or
100 such objects, it would be possible to see whether the SN \tomas\
$\mu$ discrepancy is simply an outlier, or an indication of a more
pernicious systematic error.  Any cluster that has been vetted by
pencil-beam magnification tests using lensed \SNIa\ will be able to
provide a more reliable measure of the magnifications for very high
redshift objects.  Similarly, a broad sample of SNe like \tomas\ would
help to define the preferred lens modeling methodology by highlighting
any models that consistently perform well in \SNIa\ lensing tests.
The ongoing FrontierSN program will discover and follow any more
highly magnified SNe that appear behind the Frontier Field
clusters. Unfortunately, the HFF survey is not designed with high-z
transient discovery as a primary science goal, so the FrontierSN
effort will likely add no more than 1-3 new lensed \SNIa. Further
imaging of strong-lensing clusters with \HST\ or the James Webb Space
Telescope (JWST) could enable a larger sample to be collected,
especially if the filters and cadence are optimized for detection
of \SNIa\ at $z>1$.  Massive clusters such as Abell 2744 will continue
to be attractive as cosmic telescopes, allowing the next generation of
telescopes to reach the faintest objects in the very early universe.
The puzzling bias revealed by SN \tomas\ supports a concerted effort
to improve these lenses with \change{further examination of lensing
systematics through simulations and collection of} a larger sample of
magnified SNe.

>>>>

\change{
One of the key values in observing standard candles behind
gravitational lenses is \changeG{in addressing the problem of the
mass-sheet degeneracy} \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of uniform mass in front of or behind the
lens, without disturbing the primary observable quantities.  For
example, take a lens model with a given surface mass density $\kappa$,
and then transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.
}

@@

\changeG{
Spectra of Type I SNe (including all sub-classes Ia, Ib and Ic) are
dominated by broad absorption features, which cannot in general be
used to directly extract a spectroscopic redshift \citep[see
e.g.][]{Filippenko:1997}.  As described below, we fit template spectra
to the SN \tomas\ data in two steps.}  First we determine a spectral
classification -- and get a preliminary estimate of the redshift and
age -- using the SuperNova IDentification (SNID)
software \citep{Blondin:2007}.  Second, we refine the redshift and age
measurement using a custom Type Ia spectral template matching program.

@@

As in \P14, we derive a distance modulus\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.} \changeG{while ignoring the
effects of gravitational lensing, so that we can later use it to
derive a direct measurment of the magnification.}  From the SALT2 fit
we get

@@

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  \change{Six of the tested models
(CATS, GLAFIC, Zitrin-NFW, Bradac(v2), Merten, Diego) report a median
value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and two of the free-form models (Williams,
Diego) report uncertainties that overlap the central $\mu$ as measured
from both of the SN light curve fitters.}  Furthermore, the scatter
amongst models provides a reasonable estimate of the magnification
uncertainty.

@@

\change{
\subsection{Misidentification of System 3}
\label{sec:MisidentificationOfSystem3}
}
\change{
It is important to note that in regards to the impact on this
particular sight-line, not all strong-lensing constraints are equal.
An error in the association of multiple images or in the redshift of a
multiply-imaged galaxy would have a more significant impact if that
system has a projected position close to our SN
sight-line \changeG{\citep[see][for a quantitative assessment of this
effect]{Johnson:2014}.}  It happens that there is one such system, for
which the position of a multiple image near SN \tomas\ is presently
disputed.  }

@@

We have speculated on the origin of this systematic bias,
evaluating six possible causes:

@@

\changeG{
At least for this particular sight-line, the free-form models perform
better in the sense that they deliver more realistic uncertainties.
This may indicate that the free-form approach in general is more
successful in capturing the uncertainty from biases inherent to the
mass modeling exercise, such as the mass sheet degeneracy and more
complex ``approximate
invariances'' \citep{Liesenborgs:2012,Schneider:2014}.  }

@@

Increasing the sample of SNe behind clusters \changeG{like those in
the HFF program -- with rich lensing constraints and deep imaging -- }
would allow this test to be repeated and refined.  With 10 or 100 such
objects, it would be possible to see whether the SN \tomas\ $\mu$
discrepancy is simply an outlier, or an indication of a more
pernicious systematic error.  Any cluster that has been vetted by
pencil-beam magnification tests using lensed \SNIa\ will be able to
provide a more reliable measure of the magnifications for very high
redshift objects.  Similarly, a broad sample of SNe like \tomas\ would
help to define the preferred lens modeling methodology by highlighting
any models that consistently perform well in \SNIa\ lensing tests.
The ongoing FrontierSN program will discover and follow any more
highly magnified SNe that appear behind the Frontier Field
clusters. Unfortunately, the HFF survey is not designed with high-z
transient discovery as a primary science goal, so the FrontierSN
effort will likely add no more than 1-3 new lensed \SNIa. Further
imaging of strong-lensing clusters with \HST\ or the James Webb Space
Telescope (JWST) could enable a larger sample to be collected,
especially if the filters and cadence are optimized for detection
of \SNIa\ at $z>1$.  Massive clusters such as Abell 2744 will continue
to be attractive as cosmic telescopes, allowing the next generation of
telescopes to reach the faintest objects in the very early universe.
The puzzling bias revealed by SN \tomas\ supports a concerted effort
to improve these lenses with \change{further examination of lensing
systematics through simulations and collection of} a larger sample of
magnified SNe.
